---
title: "BIOL4408 Marine Ecology: lobster density 1."
author: "TimLanglois"
date: "16/02/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#### Pacakges we will need

library(googlesheets4) #to read gsheet
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files


# How to install packages

#install.packages("googlesheets4")
#install.packages("tidyr")
#install.packages("dplyr")
#install.packages("readr")
#install.packages("here")
#install.packages("ggplot2")
#install.packages("plyr")

```





## 1. Import lobster density data from a google sheet

We have to install two packages that are not pre-installed in ecocloud

```{r}
# install.packages("googlesheets4")

# install.packages("here")
```




Next we load some librarys
```{r libs1, message=FALSE, warning=FALSE}
library(googlesheets4) #to read gsheet
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files
```

It is useful to set a study name
```{r name, message=FALSE, warning=FALSE}
study<-"lobster.density"
```

Read in the data from the google sheet and check it
```{r read, message=T, warning=FALSE}
#We need the URL for the google sheet
url <- "https://docs.google.com/spreadsheets/d/1Wqn7m2jopx11n5fdl9MHZAujRVkjBusdmIHq_gMhf0A"

# Then we can read it in
options(httr_oob_default=TRUE) # to allow access to googlesheets

dat<-read_sheet(url, sheet = "lobster.density")%>%
  as_tibble()%>%
  select(-c('longitude','latitude','time','way.point','gps','group'))%>%
  glimpse()
```

Check we have all ther years of data!

```{r}
summary(dat$year)
```


Now we need to save a copy of the data.

There are several ways to setting the directory where we will read data from or write data to.

We are going to use the here() function, which creates a shortcut to your location.

```{r , message=FALSE, warning=FALSE}


here()
```

As long as the names on the folders are consistent this function will enable us to work across computers and operating systems.


If you are using an ecocloud server - we will have to add folder names to the here() function - but this should work

#here("workspace","Template-lobster-density","Data")

Let's create a "Data" directory then use here() to make a shortcut to that "Data" directory.
```{r here, message=FALSE, warning=FALSE}
dir.create(file.path(here(), "Data")) #create Data folder

data.dir <- here("Data")

#or for ecocloud
#dir.create(file.path(here(), "workspace","Template-lobster-density","Data")) #create Data folder

#data.dir <- here("workspace","Template-lobster-density","Data")
```


Now to write the data we have imported from the googlesheet.
We will append the study name.
```{r write, message=TRUE, warning=TRUE}
setwd(data.dir)#this is our shortcut using here()
dir()

write_csv(dat,paste(study,"gsheet","csv",sep = "."))

```



Go back to [Analysis-lobster-density](https://github.com/UWA-BIOL4408-Marine-Ecology/Analysis-lobster-density/blob/master/README.md)

Go forward to [2_lobster-density_check-data](https://github.com/UWA-BIOL4408-Marine-Ecology/Analysis-lobster-density/blob/master/2_lobster-density_check-data.md)