---
title: "BIOL4408 Marine Ecology: Lobster density 3."
author: "TimLanglois"
date: "09/01/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 3. Basic plots to check the data


Load extra librarys
```{r libs2, message=FALSE, warning=FALSE}
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files

library(forcats) #to transform catagorical data
library(ggplot2) #to plot data
library(RCurl) #needed to download data from GitHub

```


Set a study name
```{r name, message=FALSE, warning=FALSE}
study<-"lobster.density"
```


Use here() to make a shortcut to the "Data" directory.
```{r here, message=FALSE, warning=FALSE}
here()
data.dir <- here("Data")

data.dir
```


## Read in the data and glimpse the data.
glimpse() shows the variable names, formats, head of the data.

I have also shown you the message from read_csv() - which alerts you to the format that each column of the data was read in as.
read_csv() decides on this format by previewing the first 1000 rows of data.

This is very important to check and you can also use this column specification to stipluate/force the formatting of each column as the data is read in by the function.

```{r read, message=TRUE, warning=FALSE}
setwd(data.dir)#this is out shortcut using here()
dir()

dat<-read_csv("lobster.density.csv")%>%
  glimpse()

```


## OR read from github and write to data folder
```{r message=FALSE, warning=FALSE}

dat.git<-read.csv(text=getURL("https://raw.githubusercontent.com/UWA-BIOL4408-Marine-Ecology/Analysis-lobster-density/master/Data/lobster.density.csv"))%>%
  glimpse()


setwd(data.dir) #set the directory
dir() #look in the directory

# Write dat using study name
write_csv(dat,paste(study,"csv",sep = "."))

```



## Basic plots to check out the data


Scatterplot
```{r}
ggplot(data=dat, aes(x=status, y=count)) + 
  geom_point()

```


Scatterplot with jitter and alpha (great for checking data)
```{r}
ggplot(dat, aes(x=status, y=count)) + 
  geom_point(position = position_jitter(),alpha = 1/4) #alpha gives transparency
```


Scatterplot with jitter and alpha and with one factor facetted
```{r}
ggplot(dat, aes(x=status, y=count)) + 
  geom_point(position = position_jitter(),alpha = 1/4)+
  facet_grid(.~year) #facet by factor
```


Scatterplot with jitter and alpha and with two factors facetted
```{r}
ggplot(dat, aes(x=status, y=count, colour=status)) + 
  geom_point(position = position_jitter(width = 0.1, h = 0),alpha = 1/4)+
  facet_grid(size.class~year)
```


Boxplot with facet by two factors with outliers
```{r}
ggplot(dat, aes(x=status, y=count)) + 
  geom_boxplot()+
  facet_grid(size.class~sanctuary)
```


Boxplot with facet by two factors without outliers
```{r}
ggplot(dat, aes(x=status, y=count)) + 
  geom_boxplot(outlier.shape = NA)+
  facet_grid(size.class~sanctuary)
```

Box plots provide a summary of your data:
- median (Q2/50th Percentile): the middle value of the dataset.
- first quartile (Q1/25th Percentile): the middle number between the smallest number (not the “minimum”) and the median of the dataset.
- third quartile (Q3/75th Percentile): the middle value between the median and the highest value (not the “maximum”) of the dataset.
- interquartile range (IQR): 25th to the 75th percentile.
- “maximum” whisker: Q3 + 1.5*IQR
- “minimum” whisker: Q1 -1.5*IQR
- outliers

Boxplots are useful to tell you about your outliers and what their values are. They can also tell you if your data is symmetrical, how tightly your data is grouped, and if and how your data is skewed.

Our lobster data is heavily skewed to zero values. Box plots are not so useful to display such data, but it is important for us to see this zero-richness and think about how to account for this in our data analysis.

Boxplots do not indicate where the mean is. Given that most statistical tests are comparing means - we should add a mean to see where it is.



Boxplot with facet by two factors, scatterplot with alpha and mean.

```{r}
ggplot(dat, aes(x=status, y=count)) + 
  geom_boxplot(outlier.shape = NA)+
  geom_point(position = position_jitter(width = 0.3, h = 0),alpha = 1/4, size=1)+
  stat_summary(fun.y=mean, geom="point", shape=2, size=4, colour="red")+ #adds mean
  facet_grid(size.class~sanctuary)
```

The means are close to zero!
We will make some alternative plots, better suited to showing zeros in the next script.


Boxplots are not good for displaying count data, especially zero-rich count data.
Boxplots are much better suited to display continous environmental data such as our estimates of habitat complexity.

Our complexity data is only semi-continous and really ordered categorical data, but the boxplot can show us the patterns in this data.

```{r}
ggplot(dat, aes(x=status, y=complexity)) + 
  geom_boxplot(outlier.shape = NA)+
  geom_point(position = position_jitter(width = 0.3, h = 0),alpha = 1/4, size=1)+
  stat_summary(fun.y=mean, geom="point", shape=2, size=4, colour="red")+ #adds mean
  facet_grid(size.class~sanctuary)
```


And what does our depth data from 2020 look like?
```{r}
ggplot(dat%>%filter(year==2020), aes(x=status, y=depth)) + 
  geom_boxplot(outlier.shape = NA)+
  geom_point(position = position_jitter(width = 0.3, h = 0),alpha = 1/4, size=1)+
  stat_summary(fun.y=mean, geom="point", shape=2, size=4, colour="red")+ #adds mean
  facet_grid(size.class~sanctuary)
```


Go back to [Analysis-lobster-density](https://github.com/UWA-BIOL4408-Marine-Ecology/Analysis-lobster-density/blob/master/README.md)

Go forward to [4_lobster-density_advanced-plots](https://github.com/UWA-BIOL4408-Marine-Ecology/Analysis-lobster-density/blob/master/4_lobster-density_advanced-plots.md)
