---
title: "BIOL4408 Marine Ecology: Lobster density 4."
author: "TimLanglois"
date: "09/01/2020"
output: github_document
---

```{r setup, include=FALSE}

library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```


# 4. Advanced plots

For a discussion on good practice for plotting - see here: ********

Rember that is it generally accepted that you should only plot by significant/important factors/relationships. So your final plotting should reflect what you find to be significant/important in your data analysis.

Here I provide exmaples of how to plot the potentially significant/important factors.



Load extra librarys
```{r libs2, message=FALSE, warning=FALSE}
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files

library(forcats) #to transform catagorical data
library(ggplot2) #to plot data
```


Set a study name
```{r name, message=FALSE, warning=FALSE}
study<-"lobster.density"
```


Create functions to plotting SE
```{r , message=FALSE, warning=FALSE}
se <- function(x) sd(x) / sqrt(length(x))
se.min <- function(x) (mean(x)) - se(x) #to make SE min.
se.max <- function(x) (mean(x)) + se(x) #to make SE max.
```


Use here() to make a shortcut to the "Data" directory.
```{r here, message=FALSE, warning=FALSE}
data.dir <- here("Data")
```


Create a 'Plots' folder if you don't already have one and create a shortcut to it
```{r , message=FALSE, warning=FALSE}
dir.create(file.path(here(), "Plots")) #create Plots folder

plots.dir=here("Analysis_Lobster_density","Plots")
```


Read in data for plotting
```{r read, message=TRUE, warning=FALSE}
setwd(data.dir)#this is out shortcut using here()
dir()

dat<-read_csv("lobster.density.csv")%>%
  glimpse()

```


## Make mean and SE plot

Make a mean +/-SE plot with a title.
Here we use stat_summary() and the functions we made above.
```{r}

ggplot(dat%>%filter(size.class=="legal"), aes(x=status, y=count,fill=status)) + 
  stat_summary(fun.y=mean, geom="bar") + #add bar at mean
  stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1) #add error bars

```


## Save a plot
Here we use ggsave(), you can specify the format of the plot, it's size and many other settings. By default it will use the last plot made and set the size of the plot to the size of the plotting window. 
```{r}

ggsave(here("Plots","status.barplot.png"))

```


When you make a plot you can create a plot object and then call this in the ggsave() - we have also added a title
```{r message=TRUE, warning=TRUE}
status.barplot <- ggplot(dat%>%filter(size.class=="legal"), aes(x=status, y=count,fill=status)) + 
  stat_summary(fun.y=mean, geom="bar") + #add bar at mean
  stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1) + #add error bars
  ggtitle("Legal lobster") #add a title
status.barplot

```

Here we save the plot base on its name.
This makes it simpler when you are making mulitple plots in a script and want to save them all in one go at the end of the script.
We can also sett the size of the final plot
```{r}
ggsave(here("Plots","status.barplot2.png"),status.barplot)


ggsave(here("Plots","status.barplot2.png"),status.barplot,
       width = 15, height = 8,units = "cm") 

```



## What's wrong with this code?

ggbarplot.status.size<-ggplot(dat, aes(x=status, y=count,fill=status)) +
stat_summary(fun.y=mean, geom="bar") +
stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1)+
facet_grid(size.class~)

ggbarplot.status.size



# Very smart plot using Themes

Plotting in R is very customisable. And easy way to style the asethics of your plots is to use different default themes, e.g. theme_bw().

In addition you can develop your own custom theme to augment the default themes.
Use theme_get() to see available options.


## Themes

An example theme to simplify our plots. Use # to experiment turning on and off different options.
```{r}
Theme1 <-
  theme(
    panel.grid.major = element_blank(), #turns off the major grid lines
    panel.grid.minor = element_blank(), #turns off the minor grid lines
    
    text=element_text(size=14), #sets all text to 14
    
    axis.title.x=element_text(vjust=0.3, size=15),
    axis.title.y=element_text(vjust=0.6, angle=90, size=15),
    axis.text.y=element_text(size=14),
    axis.line.x=element_line(colour="black", size=0.5,linetype='solid'),
    axis.line.y=element_line(colour="black", size=0.5,linetype='solid'),
    
    legend.title = element_blank(), #turns off the legend title
    legend.background = element_blank(), #turns off the legend outline
    legend.key = element_blank(), # turns off the rectangle around symbols in the legend
    legend.text = element_text(size=15),
    legend.position = "top",
    
    strip.text.y = element_text(size = 14,angle = 0), #sets size and angle of text in facet strip
    strip.background = element_blank()) #turns off the facet strip box
```


Mean and +/-SE plot with theme_bw() and custom theme
```{r}

status.year.sanctuary<-ggplot(dat%>%filter(size.class=="legal"), aes(x=status, y=count,fill=status)) + 
  stat_summary(fun.y=mean, geom="bar", colour="black") +
  stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1) +
  # Labels
  xlab("Marine sanctuary")+
  ylab(bquote('Density of legal rock lobster (no. /10 m'^2*')'))+
  ggtitle("Density of legal lobster")+
  # Apperance
  theme_bw()+
  Theme1+
  facet_grid(year~sanctuary)
status.year.sanctuary

```


Save the plot - change the size to fit all the text in.
```{r echo=TRUE}
ggsave(here("Plots","status.barplot.png"),status.year.sanctuary,
       width = 15, height = 15,units = "cm")

```





## Keep on plotting
If you want to learn more just Google on StackOverflow and other help forums.
Also see http://www.cookbook-r.com/Graphs/ for some handy plotting and formatting help.


# Some more interesting plots


All years - Year x Status
```{r echo=TRUE}
ggplot(dat%>%filter(size.class=="legal"), aes(x=year, y=count,colour=status)) + 
  stat_summary(fun.y=mean, geom="point") +
  stat_summary(fun.y=mean, geom="line") +
  stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1)+
  theme_bw()
```


All years - complexity x Sanctuary x Status
```{r}
ggplot(dat%>%filter(size.class=="legal"), aes(x=complexity, y=count,colour=status)) + 
  geom_smooth(method=lm, size=0.5,se=F)+
  theme_bw()+
  facet_grid(sanctuary~.)
```

2019 - Sanctuary x Status
```{r}
ggplot(dat%>%filter(size.class=="legal"&year==2019),aes(x=status, y=count,fill=status)) +
    stat_summary(fun.y=mean, geom="bar", colour="black") +
    stat_summary(fun.ymin = se.min, fun.ymax = se.max, geom = "errorbar", width = 0.1) +
    facet_grid(.~sanctuary)
```


2019 -  Complexity x Sanctuary
```{r}
 ggplot(dat%>%filter(size.class=="legal"&year==2019), aes(x=complexity, y=count,colour=sanctuary)) + 
    geom_smooth(method=lm, size=0.5,se=F)+
    coord_cartesian(ylim = c(0, 3), expand = FALSE)
```


## A note on geom_smooth()

Geom_smooth() will calculate SE around a fitted model dependent on the method argument.
This SE is based a simplest model, as specified in the method argument, and is unlikley to be appropriate for the sampling design or data analysis model you have used.

In the case of this example lobster density data, the sampling design involes multiple locations and sites, which may have been analysed with a nested and/or mixed model. Therefore any SE calcualted by the geom_smooth() function will not be correct/appropriate to display.
  
  
2019 -  Complexity x Sanctuary with an incorrect SE
```{r}
 ggplot(dat%>%filter(size.class=="legal"&year==2019), aes(x=complexity, y=count,colour=sanctuary)) + 
    geom_smooth(method=lm, size=0.5,se=T)+
    coord_cartesian(ylim = c(0, 3), expand = FALSE)+
  ggtitle("incorrect SE")
```

In this case, the appropriate way to generate a SE would be to model the data using glmer() or equivalent and then use predict() to predict the mean and SE of the fitted model.

No one worries about this with most ANOVA analyses :)


Go to [5_lobster-density_format-for-PRIMER](https://github.com/UWA-BIOL4407-Marine-Ecology/Analysis-lobster-density/blob/234351d599d9e8c931c3bdf610e34f7980c6b273/5_lobster-density_format-for-PRIMER.md)

Go back to [Analysis-lobster-density](https://github.com/UWA-BIOL4407-Marine-Ecology/Analysis-lobster-density/blob/234351d599d9e8c931c3bdf610e34f7980c6b273/README.md)



